Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-07-09 2016-Weak-FH-Bisim/WeakBisim.tex
Tentatives sur lemma 6

\begin{lemma}[Weak OT decomposition]\label{lem-decomposeWOT} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a pLTS but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.  
		 Then there exist $n$, $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\beta_p$, $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
\text{\qquad and \qquad}\gamma_{j_0}=[\beta_0..\beta_n]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mybigdotcup_{p=0}^n \gamma_{p j} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred = \Pred\,'
		\land \!\!\bigwedge_{p=0}^n((\alpha_p\subst{\bigotimes_{i=p-1}^{0}\Post_i)} = \beta_p\land (\Pred_p\subst{\bigotimes_{i=p-1}^{0}\Post_i)} )
\\
\text{and~}
\Post=\Post\,' \uplus \bigotimes_{p=n}^0
		\Post_p  
		\end{mathpar} 
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}

@@

Let us consider the sequence of $(n_1+n_2+3)$  actions $\alpha_p$ some of them may be non-observable (are $\tau$ transition). By considering $\tau$  actions we split the sequence of actions into $n$ subsequences, such that each subsequence is a sequence of actions containing  only one observable action, and possibly many $\tau$ ones.  We can decompose each of the $n$ subsequences; for $k\in [1..n]$ the position of the $k^{\text{th}}$ visible action is is $n_k$, $k\in [1..n-1]$ for  $n'_k$ is any index between $n_k$ and $n_{k+1}$,  $n'_0=0$ and $n'_n=n_1+n_2+3$. We obtain $n$ subsequences  made of the following OTs, for all $k\in [1..n]$ :\\
$\forall p\! \in\! [n'_{k-1}..(n_k\!-\!1)] Q\models{\openrule
			{
				\beta_{pj}^{j\in J\cap\Holes(Q)}, {{\Pred''_{p}}},  
				{{\Post''_{p}}}}
				{\ostate{v_{pl}^{l\in L_Q}} \OTarrow {\tau}
				\ostate{v_{(p+1)l}^{l\in L_Q}}}
				}%\]
,$ 
$Q\models{\openrule
			{
				\beta_{n_k j}^{j\in J\cap\Holes(Q)}, {{\Pred''_{n_k}}},  
				{{\Post''_{n_k}}}}
				{\ostate{v_{n_kl}^{l\in L_Q}} \OTarrow {\alpha_{{n_k}}}
				\ostate{v_{(n_k+1)l}^{l\in L_Q}}}
				}%\]
$
and
$\forall p \in [(n_k\!+\!1)..(n_k'\!-\!1)]\,\, Q\models{\openrule
			{
				\beta_{pj}^{j\in J\cap\Holes(Q)}, {{\Pred''_{p}}},  
				{{\Post''_{p}}}}
				{\ostate{v_{pl}^{l\in L_Q}} \OTarrow {\tau}
				\ostate{v_{(p+1)l}^{l\in L_Q}}}
				}%\]
$\\ 
Thereafter, by Lemma \ref{lem-rel-OT-WOT} we can deduce the following weak open transition:
\begin{mathpar}
Q\models{\openrule
			{
			(\gamma_{kj})^{j\in J\cap\Holes(Q)},  {{\Pred_k}},  
				{{\Post_k}}}
			{\ostate{v_{kl}^{l\in L_Q}} \OTWeakarrow {\alpha'_{k}}
				\ostate{(v_{kl}')^{l\in L_Q}}}
		}%\]
\end{mathpar}
$\alpha_k'=\alpha_{n_k}\displaystyle{\subst{\bigotimes_{j=n_k-1}^{n'_{k-1}}\!\!\Post''_{j}}}$\\
$\forall l \in L_Q. v_{kl}=v_{(n'_{k-1})l} \wedge v_{kl}'=v_{(n'_k-1)l}$\\
$\displaystyle{\gamma_{kj}^{j\in J\cap\Holes(Q)} }=\displaystyle{\mybigdotcup_{i=n'_{k-1}}^{(n_k\!-\!1)} \vis{{\beta_{ij}^{j\in J\cap\Holes(Q)}}\subst{\bigotimes_{l=i-1}^{n'_{k-1}}\!\!\!\Post''_{l}}}\dotcup \vis{{\beta_{n_k j}^{j\in J\cap\Holes(Q)}}\subst{\bigotimes_{l=n_k-1}^{n'_{k-1}}\!\!\!\Post''_{l}}}} \dotcup\\
~~\qquad \mybigdotcup_{i=n_k+1}^{n_k'-1}
\vis{{\beta_{ij}^{j\in J\cap\Holes(Q)}} \subst{\bigotimes_{l=i-1}^{n_k\!+\!1}\!\Post''_{l}\shortotimes\Post''_{n_k}\shortotimes\!\!\bigotimes_{l=n_k-1}^{n'_{k-1}}\!\!\!\Post''_{l}} }\\
%
\Pred_k=\bigwedge_{i=n'_{k-1}}^{n_k\!-\!1}\!\!\Pred''_{i}\subst{\bigotimes_{j=i-1}^{n'_{k-1}}\!\!\Post_{j}}\land\Pred''_{n_k} \subst{\bigotimes_{j=n_k-1}^{n'_{k-1}}\!\!\Post''_{j}}\land\\ 
~\qquad\quad\bigwedge_{i=n_k+1}^{n_k'-1}\!\!\!\Pred''_{i}\subst{\bigotimes_{j=i-1}^{n'_{k-1}}\!\!\!\Post''_{j}\shortotimes\Post''_{n_k}\shortotimes\!\!\bigotimes_{j=n_k-1}^{n'_{k-1}}\!\!\!\Post''_{j}}
\\
\Post_k= \bigotimes_{j=n_k'-1}^{n'_{k-1}}\!\!\Post''_{j}$\\

Also note that for all $k\in[1..n-1]$, $v_{kl}'=v_{(k+1)l}$, and $v_{0l}=s_{0l}=s_l$ and $v'_{nl}=v_{(n_1+n_2+3) l}=u_{n_2 l} = s'_l$.

%In this case, the number of weak open transitions
%might not be exactly the number of open transitions that take part in the reduction but something smaller, thus the resulting set of weak open transitions could also be smaller.\\
Note that \[\alpha_{n_k}\displaystyle{\subst{\bigotimes_{j=n_k-1}^{0}\!\!\Post''_{j}}} =
\alpha_{n_k}\displaystyle{\subst{\bigotimes_{j=n_k-1}^{n'_{k-1}}\!\!\Post''_{j}\otimes\bigotimes_{j=n'_{k-1}-1}^{0}\!\!\Post''_{j}}} = \alpha_k'\displaystyle{\subst{\bigotimes_{j=n'_{k-1}-1}^{0}\!\!\Post''_{j}}} = 
\alpha_k'\displaystyle{\subst{\bigotimes_{j=k-1}^{0}\!\!\Post_{j}}}\]
We now have (the $\alpha'_p$ correspond to the observable $\alpha_k$ actions): \TODO{add ref to eq 1 page above?}
\begin{eqnarray*}
\Pred_\beta &\iff& \gamma_{j_0}=
\vis{[\alpha_{p}\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{j}}|p\in[0..n_1+n_2+2]]}\\
%&\iff& \gamma_{j_0}=
%\vis{[\alpha'_{p}\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{j}}|p\in[1..n]]}\\
&\iff& \gamma_{j_0}=
[\alpha'_{p}\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{j}}|p\in[1..n]]
\end{eqnarray*}

We need now to show that this set of WOT verify the lemma.

>>>>

\begin{lemma}[Weak OT decomposition]\label{lem-decomposeWOT} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a pLTS but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.  
		 Then there exist $n$, $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\beta_p$, $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
\text{\qquad and \qquad}\gamma_{j_0}=[\beta_0..\beta_n]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{and~~}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) \qquad
 \text{and\qquad} \gamma_j^{j\in J\cap\Holes(Q)}= \mybigdotcup_{p=0}^n \gamma_{p j}^{j\in J_p} \\
 \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred = \Pred\,'
		\land \!\!\bigwedge_{p=0}^n((\alpha_p\subst{\bigotimes_{i=p-1}^{0}\Post_i)} = \beta_p\land (\Pred_p\subst{\bigotimes_{i=p-1}^{0}\Post_i)} )
\\
\text{and~}
\Post=\Post\,' \uplus \bigotimes_{p=n}^0
		\Post_p  
		\end{mathpar} 
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}

@@

Let us consider the sequence of $(n_1+n_2+3)$  actions $\alpha_p$ some of them may be non-observable (are $\tau$ transition). By considering $\tau$  actions we split the sequence of actions into $n+1$ subsequences, such that each subsequence is a sequence of actions containing  only one observable action, and possibly many $\tau$ ones.  We can decompose each of the $n+1$ subsequences; for $k\in [0..n]$ the position of the $k^{\text{th}}$ visible action is is $n_k$, $k\in [1..n-1]$ for  $n'_k$ is any index between $n_k$ and $n_{k+1}$,  $n'_0=0$ and $n'_{n+1}=n_1+n_2+3$. We obtain $n+1$ subsequences  made of the following OTs, for all $k\in [0..n]$ :\\
$\forall p\! \in\! [n'_{k}..(n_k\!-\!1)] Q\models{\openrule
			{
				\beta_{pj}^{j\in J\cap\Holes(Q)}, {{\Pred''_{p}}},  
				{{\Post''_{p}}}}
				{\ostate{v_{pl}^{l\in L_Q}} \OTarrow {\tau}
				\ostate{v_{(p+1)l}^{l\in L_Q}}}
				}%\]
,$ 
$Q\models{\openrule
			{
				\beta_{n_k j}^{j\in J\cap\Holes(Q)}, {{\Pred''_{n_k}}},  
				{{\Post''_{n_k}}}}
				{\ostate{v_{n_kl}^{l\in L_Q}} \OTarrow {\alpha_{{n_k}}}
				\ostate{v_{(n_k+1)l}^{l\in L_Q}}}
				}%\]
$
and
$\forall p \in [(n_k\!+\!1)..(n'_{k+1}\!-\!1)]\,\, Q\models{\openrule
			{
				\beta_{pj}^{j\in J\cap\Holes(Q)}, {{\Pred''_{p}}},  
				{{\Post''_{p}}}}
				{\ostate{v_{pl}^{l\in L_Q}} \OTarrow {\tau}
				\ostate{v_{(p+1)l}^{l\in L_Q}}}
				}%\]
$\\ 
Thereafter, by Lemma \ref{lem-rel-OT-WOT} we can deduce the following weak open transition:
\begin{mathpar}
Q\models{\openrule
			{
			(\gamma_{kj})^{j\in J\cap\Holes(Q)},  {{\Pred_k}},  
				{{\Post_k}}}
			{\ostate{v_{kl}^{l\in L_Q}} \OTWeakarrow {\alpha'_{k}}
				\ostate{(v_{kl}')^{l\in L_Q}}}
		}%\]
\end{mathpar}
$\alpha_k'=\alpha_{n_k}\displaystyle{\subst{\bigotimes_{j=n_k-1}^{n'_{k}}\!\!\Post''_{j}}}$\\
$\forall l \in L_Q. v_{kl}=v_{(n'_{k})l} \wedge v_{kl}'=v_{(n'_k)l}$\\
$\displaystyle{\gamma_{kj}^{j\in J\cap\Holes(Q)} }=\displaystyle{\mybigdotcup_{i=n'_{k}}^{(n_k\!-\!1)} \vis{{\beta_{ij}^{j\in J\cap\Holes(Q)}}\subst{\bigotimes_{l=i-1}^{n'_{k}}\!\!\!\Post''_{l}}}\dotcup \vis{{\beta_{n_k j}^{j\in J\cap\Holes(Q)}}\subst{\bigotimes_{l=n_k-1}^{n'_{k}}\!\!\!\Post''_{l}}}} \dotcup\\
~~\qquad \mybigdotcup_{i=n_k+1}^{n'_{k+1}-1}
\vis{{\beta_{ij}^{j\in J\cap\Holes(Q)}} \subst{\bigotimes_{l=i-1}^{n_k\!+\!1}\!\Post''_{l}\shortotimes\Post''_{n_k}\shortotimes\!\!\bigotimes_{l=n_k-1}^{n'_{k}}\!\!\!\Post''_{l}} }\\
%
\Pred_k=\bigwedge_{i=n'_{k}}^{n_k\!-\!1}\!\!\Pred''_{i}\subst{\bigotimes_{j=i-1}^{n'_{k}}\!\!\Post_{j}}\land\Pred''_{n_k} \subst{\bigotimes_{j=n_k-1}^{n'_{k}}\!\!\Post''_{j}}\land\\ 
~\qquad\quad\bigwedge_{i=n_k+1}^{n'_{k+1}-1}\!\!\!\Pred''_{i}\subst{\bigotimes_{j=i-1}^{n'_{k}}\!\!\!\Post''_{j}\shortotimes\Post''_{n_k}\shortotimes\!\!\bigotimes_{j=n_k-1}^{n'_{k}}\!\!\!\Post''_{j}}
\\
\Post_k= \bigotimes_{j=n'_{k+1}-1}^{n'_{k}}\!\!\Post''_{j}$\\

Also note that for all $k\in[0..n-1]$, $v_{kl}'=v_{(k+1)l}$, and $v_{0l}=s_{0l}=s_l$ and $v'_{nl}=v_{(n_1+n_2+3) l}=u_{n_2 l} = s'_l$.

%In this case, the number of weak open transitions
%might not be exactly the number of open transitions that take part in the reduction but something smaller, thus the resulting set of weak open transitions could also be smaller.\\
Note that \[\alpha_{n_k}\displaystyle{\subst{\bigotimes_{j=n_k-1}^{0}\!\!\Post''_{j}}} =
\alpha_{n_k}\displaystyle{\subst{\bigotimes_{j=n_k-1}^{n'_{k}}\!\!\Post''_{j}\otimes\bigotimes_{j=n'_{k}-1}^{0}\!\!\Post''_{j}}} = \alpha_k'\displaystyle{\subst{\bigotimes_{j=n'_{k}-1}^{0}\!\!\Post''_{j}}} = 
\alpha_k'\displaystyle{\subst{\bigotimes_{j=k-1}^{0}\!\!\Post_{j}}}\]
We now have (the $\alpha'_p$ correspond to the observable $\alpha_k$ actions): \TODO{add ref to eq 1 page above?}
\begin{eqnarray*}
\Pred_\beta &\iff& \gamma_{j_0}=
\vis{[\alpha_{p}\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{j}}|p\in[0..n_1+n_2+2]]}\\
%&\iff& \gamma_{j_0}=
%\vis{[\alpha'_{p}\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{j}}|p\in[1..n]]}\\
&\iff& \gamma_{j_0}=
[\alpha'_{p}\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{j}}|p\in[1..n]]
\end{eqnarray*}

We need now to show that this set of WOT verify the conditions of the lemma, i.e. it is a set of WOT of the form:
\[	Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}\]
With
\[  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) \qquad\text{trivial}\]
\[\text{and\qquad} \gamma_j^{j\in J\cap\Holes(Q)}= \mybigdotcup_{p=0}^n \gamma_{p j}^{j\in J_p}\]
Indeed we have:
\begin{equation*}
\begin{split}
\gamma_j^{j\in J}&=\mybigdotcup_{i=0}^{m_1} \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{0}\!\! \Post_{1j}}}\dotcup  \vis{\set{\beta_2}\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}} \dotcup \mybigdotcup_{i=0}^{m_2} \vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}} }\\
\end{split}\end{equation*}

And thus, because $\beta_{pj}^{j\in J\cap \Holes(Q)}$ are equal to  $(\beta'_{1pj})^{j\in J\cap \Holes(Q)}$, $(\beta'_{2j})^{j\in J\cap \Holes(Q)}$, and $(\beta')_{3pj}^{j\in J\cap \Holes(Q)}$ (re-indexed), and additionally $(\beta')_{1pj}^{j\in J\cap \Holes(Q)}$, $(\beta')_{2j}^{j\in J\cap \Holes(Q)}$, and $(\beta')_{3pj}^{j\in J\cap \Holes(Q)}$ are identical to the holes labels $\beta_{1kj}^{j\in J\cap \Holes(Q)}$, $\beta_{2j}^{j\in J\cap \Holes(Q)}$, and $\beta_{3kj}^{j\in J\cap \Holes(Q)}$ when $Q$ moves\footnote{more precisely, when $Q$ moves either $\beta_{1kj}^{j\in J\cap \Holes(Q)}$ is not empty  and thus $(\beta')_{1pj}^{j\in J\cap \Holes(Q)}=\beta_{1kj}^{j\in J\cap \Holes(Q)}$, or both are empty if the holes perform no action}. We can assert a similar equality on postconditions, i.e. between $\Post''_p$ and $\Post''_{1k}$, $\Post''_{2}$, $\Post''_{3k}$ where $\Post''_{1p}$ is the restriction of $\Post_{1p}$ over variables of $\Leaves(Q)$ (see initial decomposition, case 1, 2, and 3 above). Overall, we have
{\scriptsize \begin{equation*}
\begin{split}
\gamma_j^{j\in J\cap \Holes(Q)}&=\mybigdotcup_{i=0}^{m_1} \vis{\set{(\beta')_{1i}}\subst{\bigotimes_{j=i-1}^{0}\!\! \Post''_{1j}}}\dotcup  \vis{\set{\beta'_2}\subst{\bigotimes_{j=m_1}^{0}\Post''_{1j}}} \dotcup \mybigdotcup_{i=0}^{m_2} \vis{\set{\beta'_{3i}} \subst{\bigotimes_{j=i-1}^{0}\!\Post''_{3j}\shortotimes\Post''_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post''_{1j}} }\\
&=\mybigdotcup_{k=0}^n \left( \right)
\end{split}\end{equation*}
}
% \text{and\qquad} \forall j\in J_p,\,\gamma_j= \mybigdotcup_{p=0}^n \gamma_{p j} \\
% \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\{\hspace{-3ex}}\text{and~~} \Pred = \Pred\,'
%		\land \!\!\bigwedge_{p=0}^n((\alpha_p\subst{\bigotimes_{i=p-1}^{0}\Post_i)} = \beta_p\land (\Pred_p\subst{\bigotimes_{i=p-1}^{0}\Post_i)} )
